Old Data Qualified Ramsey Exploration
========================================================



```{r, cache=FALSE}
require(foreign)
getwd()
HouseData = read.dbf("~/OSI-Hedonic Project/Data/GIS2R/OldDataQualifiedRam.dbf") 
```

```{r}
options(width=160)
```

Let's start constructing our basic hedonic regression by formally creating the logged sale price variable, converting a couple variables from string to numeric, and creating some time related variables (year of sale, month of sale, etc.).
```{r, cache=FALSE}
HouseData$logPRICE = log(HouseData$PRICE)
HouseData$logSALE_VALUE = log(HouseData$SALE_VALUE)
HouseData$GARAGEsqft = as.numeric(as.character(HouseData$GARAGESQFT))
HouseData$SALE_YEAR = 1900 + as.POSIXlt(HouseData$SALE_DATE)$year
HouseData$SALE_MONTH = 1 + as.POSIXlt(HouseData$SALE_DATE)$mon
HouseData$YEARMON = 12*(HouseData$SALE_YEAR - 2009) + HouseData$SALE_MONTH # number of months since january 1, 2009
```

ORIGINAL

```{r, cache=FALSE}
OldDataExp = lm(logSALE_VALUE ~ FIN_SQ_FT + ACREAGE + YEAR_BUILT + HomeStyle + ELEMENTARY 
  + WhiteDense + fam_income  + factor(SALE_MONTH) + parkArea05*factor(SALE_YEAR) + 
  OSV05KM*factor(SALE_YEAR), 
            data = HouseData)
summary(OldDataExp)
```

ADDED STRUCTURAL CHARACTERISTICS

Interestingly, bedrooms and bathrooms are both very significant, but don't improve the model much.

```{r, cache=FALSE}
OldDataExp2 = lm(logSALE_VALUE ~ FIN_SQ_FT + BEDRMS + BATHRMS + ACREAGE + YEAR_BUILT + HomeStyle + ELEMENTARY 
  + WhiteDense + fam_income  + factor(SALE_MONTH) + parkArea05*factor(SALE_YEAR) + 
  OSV05KM*factor(SALE_YEAR), 
            data = HouseData)
summary(OldDataExp2)
```

COMPARED TO NEW RAMSEY (MORE ACCURATE?) VERSIONS OF FIELDS

Wow. Huge R-squared improvement. 

```{r, cache=FALSE}
OldDataExp3 = lm(logPRICE ~ SFLA + BEDRMS + BATHRMS + LOTSIZE + YRBLT + HomeStyle + ELEMENTARY 
  + WhiteDense + fam_income  + factor(SALE_MONTH) + parkArea05*factor(SALE_YEAR) + 
  OSV05KM*factor(SALE_YEAR), 
            data = HouseData)
summary(OldDataExp3)
```

Hmm...let's isolate where that improvement is coming from.

SAME AS v3, BUT WITHOUT NEW PRICE DATA

```{r, cache=FALSE}
OldDataExp4 = lm(logSALE_VALUE ~ SFLA + BEDRMS + BATHRMS + LOTSIZE + YRBLT + HomeStyle + ELEMENTARY 
  + WhiteDense + fam_income  + factor(SALE_MONTH) + parkArea05*factor(SALE_YEAR) + 
  OSV05KM*factor(SALE_YEAR), 
            data = HouseData)
summary(OldDataExp4)
```

Yep, the new Price Data has huge effect. Others have pretty similar records (95% plus) and don't seem to matter much, but do help a little bit. 


